---
title: "Model-agnostic: Partial Dependency Plot (PDP)"
author: "Yusef Ahsini Ouariaghli, Eva Cantín Larumbe and Pablo Díaz-Masa Valencia"
date: "2023-04-26"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. One dimensional Partial Dependence Plot.
The partial dependence plot shows the marginal effect of a feature on the predicted outcome of a previously fit model. 

## Exercise
**Apply PDP to the regression example of predicting bike rentals. Fit a random forest approximation for the prediction of bike rentals (cnt). Use the partial dependence plot to visualize the relationships the model learned. Use the slides shown in class as model.**

First and foremost, we use the perform a Random Forest to predict the number of bike rentals (cnt). 

```{r}
df <- read.csv("day.csv")
df$season <- as.factor(df$season)
df$MISTY <- ifelse(df$weathersit == 2, 1, 0)
df$RAIN <- ifelse(df$weathersit %in% c(3, 4), 1, 0)
df$temp <- df$temp * 47 - 8
df$hum <- df$hum * 100
df$windspeed <- df$windspeed * 67
df$dteday <- as.Date(df$dteday, format = "%Y-%m-%d")
df$days_since_2011 <- as.numeric(df$dteday - as.Date("2011-01-01"))

library(randomForest)
set.seed(1234)
rf <- randomForest(formula = cnt ~ workingday +  holiday + season + MISTY + RAIN + temp + hum + windspeed + days_since_2011, 
                   data = df)
```


Then, we do a Partial Dependence Plot to explain the results from the Random Forest, because it can show if the relationship between the target and a feature is linear, monotonic or more complex.

```{r}
library(pdp)
pdp1 <-partial(rf, pred.var = c("days_since_2011"), plot=TRUE)
pdp2<- partial(rf, pred.var=c("temp"), plot=TRUE)
pdp3 <- partial(rf, pred.var=c("hum"), plot=TRUE)
pdp4 <-partial(rf, pred.var=c("windspeed"), plot=TRUE)
pdp1
pdp2
pdp3
pdp4

```

